LANGuage IDentification

### 2024-12-13 paper/paper.tex
neural network and forward propagation

%    Copyright (c)  2024  João Augusto Costa Branco Marado Torres.
%    Permission is granted to copy, distribute and/or modify this document
%    under the terms of the GNU Free Documentation License, Version 1.3
%    or any later version published by the Free Software Foundation;
%    with no Invariant Sections, no Front-Cover Texts, and no Back-Cover Texts.
%    A copy of the license is included in the section entitled "GNU
%    Free Documentation License".

@@

    If that's the case the function will look like this
    \eqref{eq:linear_model_basis_f}
    \begin{equation}
        \begin{split}
            y(\mathbf{x}, \mathbf{w}) & =
            \displaystyle\sum_{j=0}^{D}\left(w_{j}\phi_{j}(x_{j})\right) = \\
            & = \mathbf{\phi}\left(\mathbf{x}\right)\cdot\mathbf{w}
        \end{split}
        \label{eq:linear_model_basis_f}
    \end{equation}

@@

    And this will be what I define as an artificial neuron \eqref{eq:neuron}.
    You can see a visual representation of it on page~\pageref{fig:neuron}

@@

    Because I couldn't find a simple way of doing anonymous functions in my
    language of choice, \href{run:../src/utils/neuron.zig}{my implementation}
    does not allow specifying basis functions nor activation functions so every
    basis function can be thought as being the identify function $
    \mathbf{\phi}(x) = x $ and the activation function can be toggled between
    the identify and sigmoid \eqref{eq:sigmoid}.

@@

    %\section{Feed-Forward Neural Network}

>>>>

%    Copyright (c)  2024  João Augusto Costa Branco Marado Torres.
%    Permission is granted to copy, distribute and/or modify this document
%    under the terms of the GNU Free Documentation License, Version 1.3
%    or any later version published by the Free Software Foundation;
%    with no Invariant Sections, no Front-Cover Texts, and no Back-Cover Texts.
%    A copy of the license is included in the section entitled "GNU
%    Free Documentation License".
%
%    You should have received a copy of the GNU Free Documentation License
%    along with this program.  If not, see <https://www.gnu.org/licenses/>

@@

    If that's the case the function will look like this
    \eqref{eq:linear_model_basis_f}.
    \begin{equation}
        \begin{split}
            y(\mathbf{x}, \mathbf{w}) & =
            \displaystyle\sum_{j=0}^{D}\left(w_{j}\phi_{j}(x_{j})\right) = \\
            & = \mathbf{\phi}\left(\mathbf{x}\right)\cdot\mathbf{w}
        \end{split}
        \label{eq:linear_model_basis_f}
    \end{equation}

@@

    And this will be what I define as an artificial neuron \eqref{eq:neuron}.
    You can see a visual representation of it on page~\pageref{fig:neuron}.

@@

    Because I couldn't find a simple way of doing anonymous functions in my
    language of choice, \href{run:../src/utils/neural.zig}{my implementation}
    does not allow specifying basis functions nor activation functions so every
    basis function can be thought as being the identify function $
    \mathbf{\phi}(x) = x $ and the activation function can be toggled between
    the identify and sigmoid \eqref{eq:sigmoid}.

@@

    \section{Feed-Forward Neural Network}

    The idea now, as explained in chapter 5.1 of \cite{bishop}, is to train
    neurons so that their weight values are good.

    Because a neuron is really just a function, a neuron can be a basis
    function of another neuron, starting a network.

    %%% create visual representation of the above

    A network is composed of layers, each with its own number of neurons.

    %%% visual representation of a layer

    Each of neuron will create a value called \textbf{activation} represented
    by $ a_{j} $ \eqref{eq:activation} where $ j $ is the neuron number (we can
    think of a layer as a column matrix of neurons), $ \mathbf{X} $ is the
    function input or the activations from the layer before after they passed
    through an activation function and $ D $ is the amount of row in $
    \mathbf{x} $ $ - 1 $.

    Let's define a way to represent a matrix row:
    $$
        A = \begin{bmatrix}a_{0}\\\vdots\\a_{N}\end{bmatrix} \in \mathbb{R}^{N
            \times M}, a_{i} \in \mathbb{R}^{M}, i \in \left\{ x \in
            \mathbb{N}: x \le N \right\}
    $$
    $$
        A_{i *} = a_{i}
    $$

    \begin{equation}
        \begin{split}
            a_{j} & = f \left( \displaystyle\sum_{i=1}^{D} \left( w_{ji}x_{i}
            \right) + (1)b_{j} \right) \\ & = f \left( \mathbf{x} \cdot W_{j *}
            \right)
        \end{split}
        \label{eq:activation}
    \end{equation}

    The last layer will be the one giving the outputs $ \mathbf{y} $.

    If a neural network has $ n $ layers, the mathematical representation for
    the output $ k $ will be something like \eqref{eq:neural_network}. $
    W^{(n)} $ are the weights of that layer. For this to work, because it's
    just matrix multiplications, we have to be aware of the dimensions of the
    matrixes. This is called \textbf{forward propagation}.

    \begin{equation}
        Y_{0} = \textbf{x}
    \end{equation}
    \begin{equation}
        Y_{n} = f_{n} \left( Y_{n - 1}^{T} \cdot w^{(n)} \right), n > 0
    \end{equation}

    \begin{equation}
        \begin{split}
            y_{k}(\mathbf{x}, \mathbf{w}) & = f_{n} \left( \displaystyle\sum_{i_{n}=0} w_{k i_{n}}^{(n)} \cdots f_{2} \left( \displaystyle\sum_{i_{2}=0} w_{i_{3}i_{2}}^{(2)} f_{1} \left( \displaystyle\sum_{i_{1}=0} w_{i_{2}i_{1}}^{(1)}x_{i_{1}} \right) \right) \cdots \right) = \\
            & = f_{n} \left( \cdots f_{2} \left( f_{1} \left( \mathbf{x} \cdot w^{(1)} \right)^{T} \cdot w^{(2)} \right)^{T} \cdots \cdot w^{(n)} \right) = \\
            & = Y_{n}
        \end{split}
        \label{eq:neural_network}
    \end{equation}

    %%% full conectivity not mandatory

    %%% visual representation of a neural network

    Layers and neural network representation can be found
    \href{run:../src/utils/neural.zig}{\texttt{/src/utils/neural.zig}}.

    %%% create x^2 sin(x) |x|
